---
title: "PM566 Final Project"
author: "Weijia Hu"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r echo=FALSE, message =FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(knitr)
library(plotly)
library(DT)
library(usmap)
```

# Introduction
In my project, I’m interested in the new cases and deaths of Covid-19. The new cases indicate how fast the virus spread and the deaths indicate how serious this disease is. 

1. Until 10/06/2020, which states contribute most to the new cases and deaths?

2. Is there any association between the number of hospitalized patients and the number of deaths for COvid-19?

3. After the National day holiday, did the number of daily new cases increase? 

# Method

- Use bar charts to assess number of cases and deaths among all states.

- Use the usmap package to draw maps in order to visually present the data.

- As for the associations between hospitalization and death & between holiday and incidence of disease, line charts were used to examine them.  

# Premilinary results 

### Summary of the data

My data are from the website of The COVID Tracking Project and about COVID-19 related data from all the states in USA. The investigators of this project would update the data each day. The date in my data set ranges from 01/22/2020 to 10/06/2020.

- There are 12138 rows and 43 columns in the raw data set. 

- The investigators assigned every state a data-quality grade based on their evaluation of the completeness of states' reporting, I removed rows whose data-quality grades are D, which I think the data are not reliable. I stored these data in the data set named "cov_d". There are 400 observations in "cov_d" data set. Let's check states in this data set in recent days, in which the public health systems of Covid-19 are considered good enough. The states are American Samoa, Northern Mariana Islands and US Virgin Islands and they are not in the mainland USA.

- After removing some observations, there are 11738 observations left.

<br>

Take a look at the data:

```{r, echo=FALSE}
cov = data.table::fread("E:/566/midterm/all-states-history.csv")
cov_d = cov[cov$dataQualityGrade=='D']
cov= cov[cov$dataQualityGrade!="D", ]
cov_1006<- cov[date=='2020-10-06']
```

```{r echo=FALSE, message =FALSE}
new_death <- cov[ ,c(1,2,4,6)]
re1<-summary(new_death)
new_case <- cov[ ,c(1,2,22,24)]
re2<-summary(new_case)
kable(re1)
kable(re2)
```


- The date variable started at 01/22/2020. The first case in USA was found on 01/21/2020. So, our data made sense. 

- Death, deathincrese, positive and positiveincrease ought to be positive numbers. However, the mininmum of deathincrease and positiveincrease are -213 and -7757, repectively. It didn't make sense. We need to remove these incorrect values. 

```{r echo=FALSE}
cov$deathIncrease[cov$deathIncrease<0] <- NA
cov$positiveIncrease[cov$positiveIncrease<0] <- NA
```

<br>

###  The bar charts

<br>


1.Total death number by states by 10/06/2020. 

- The New York state has the highest total deaths number in the USA. The state of California, New Jersey and Texas have the similar numbers and are ranked second, third and fourth respectively.

```{r echo=FALSE}
cov_1006_1 <- cov_1006 [death>5000]
cov_1006_1$state <-reorder(cov_1006_1$state, -cov_1006_1$death)
cov_1006_1 %>%
ggplot(aes(x= state, y=death, fill= state))+
  geom_bar(stat='identity')+
  geom_text(aes(label = death , vjust = -0.8, hjust = 0.5, color = state)) 
```

<br>

2. Total case number by states by 10/06/2020. 

- The California state has the highest Covid-19-related cases, which is what we all know. Interestingly, the state of New York is only ranked forth, while it has the most deaths.

```{r echo=FALSE}
  cov_1006_2 <- cov_1006 [positive>200000]
  cov_1006_2$state <-reorder(cov_1006_2$state, -cov_1006_2$positive)
  cov_1006_2 %>%
ggplot(aes(x= state, y=positive, fill= state))+
  geom_bar(stat='identity')+
  geom_text(aes(label = positive , vjust = -0.8, hjust = 0.5, color = state))
```

<br>

### The maps

<br>

1.Total death number by states by 10/06/2020. 
```{r echo=FALSE}
cov_1006_d = as.data.frame(cov_1006)
g1<-plot_usmap('states', data =cov_1006, values ='death')+
scale_fill_continuous(low = "pink", high = "darkred", name = "Total number of new cases", label = scales::comma)+
labs(title = "Total number of deaths related to Covid-19 in the USA", subtitle = "By 10/06/2020")+
  theme(legend.position = "right")
ggplotly(g1)
```
<br>

2. Total case number by states by 10/06/2020.
```{r echo=FALSE}
cov_1006_d = as.data.frame(cov_1006)
g2<-plot_usmap('states', data =cov_1006, values ='positive')+
scale_fill_continuous(low = "lightsteelblue", high = "blueviolet",name = "Total number of new cases", label = scales::comma)+
labs(title = "Total number of cases related to Covid-19 in the USA", subtitle = "By 10/06/2020")+
theme(legend.position = "right")
ggplotly(g2)
```

<br>


### Holiday and new cases

```{r echo=FALSE, message =FALSE, warning=FALSE}
cov_2week <- cov[c(4736:5477), ]
cov_2week_inci <-cov_2week[cov_2week$state==c("CA", "TX","FL","NY", "GA","IL"), ]
holi<-cov_2week_inci %>%
  ggplot(mapping = aes(x= date, y=positiveIncrease, color=state))+
  geom_line(mapping=aes(color=state))+
  facet_wrap(~state, nrow=3)
ggplotly(holi)
```

<br>

# Conclusions

- The maximum of the daily increase cases is 17820 in a certain state. The mean of th increase for every single state is 879 cases per day. The maximum of the daily increase deaths is 951 in a certain state. The mean of th increase for every single state is 24.83 deaths per day. Based on this, the Covid-19 disease could spread fast but not cause so many deaths (but we still need to wear masks and keep social distance to prevent it!).

- According to the scatter plot and its smooth line, the number of deaths and the number of hopitalization have a positive relation. Notice that the confidence intervals are getting lareger. It's because only some states have very high number of deaths.

- I selected a week before and after the National Day to analyze this question. Only in Georgia, after the National Day holiday, the daily increase cases become more. So, I don't think there is any strong relationship between them.
